Best Practices for Single Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-08-19 paper/manuscript.tex
Change TIES_MD GH repo links to TIES by @bieniekmateusz

The NEQ approaches offer some advantages over the equilibrium sampling based methods. For example, the NEQ protocol allows equilibrium simulations of each end-state to be performed independently without pre-defining a ligand network: the transitions between the ligands can be generated afterwards. This eliminates the redundant equilibrium sampling if one end state is present in multiple perturbations. Equilibrium sampling based approaches would require sampling the central ligand for each edge, while NEQ would sample it once and connect to the other ligands afterwards. The NEQ protocol also provides a simple approach in combining different structures in free energy calculations, e.g. simulations in states at $\lambda=0$ and $\lambda=1$ may be initialized with the independently generated ligand coordinates. Another example would be using true apo and holo protein structures as the end states to initialize simulations for absolute binding free energy calculations. 
%Bhati \textit{et al.} recently published two studies directly comparing the equilibrium (EQ) and NEQ approaches at unprecedented scales for both RBFE and ABFE calculations using their ensemble simulation-based TIES method~\cite{wan2023eqvsneq, bhati2025}. Further, they systematically studied the effect of various parameters involved in setting up calculations for both approaches and recommended optimal protocols in each case for industrial applications. They highlighted several limitations of the NEQ approach that should be accounted for while performing such calculations to ensure reliability of predictions.

@@

    \item \href{http://www.ties-service.org}{TIES} allows the setup of perturbation maps, hybrid topologies and structures as well as simulation-ready models for NAMD and OpenMM \url{https://github.com/UCL-CCS/TIES_MD} \cite{bieniek2023}.

@@

	\item Alchemlyb: Multipackage free energy analysis
	\url{https://github.com/alchemistry/alchemlyb}~\cite{daviddotson2020alchemistry}.
	\item pymbar: MBAR implementation, but have to roll your own analysis wrapper      
	\url{https://github.com/choderalab/pymbar} \cite{shirts2008statisticallya}.
	\item Arsenic: Standardising alchemical free energy analysis \url{https://github.com/openforcefield/Arsenic}
	\item Free Energy Workflows: Sire-specific free energy map analysis using weighted path averages \url{https://github.com/michellab/freenrgworkflows}.
    \item \href{http://www.ties-service.org}{TIES}: Perform statistical analysis of the output of alchemical ensemble simulations and report reliable predictions along with statistical uncertainties for a variety of free energy estimators \url{https://github.com/UCL-CCS/TIES_MD} \cite{bieniek2023}.
	\end{itemize}
\item[] Generally, commercial software will offer more complete pipelines in which standalone analysis applications are not necessarily needed; free and open source packages often require manual analysis but allow more flexibility and modification.
\end{itemize}

>>>>

The NEQ approaches offer some advantages over the equilibrium sampling based methods. For example, the NEQ protocol allows equilibrium simulations of each end-state to be performed independently without pre-defining a ligand network: the transitions between the ligands can be generated afterwards. This eliminates the redundant equilibrium sampling if one end state is present in multiple perturbations. Equilibrium sampling based approaches would require sampling the central ligand for each edge, while NEQ would sample it once and connect to the other ligands afterwards. The NEQ protocol also provides a simple way to combine different structures in free energy calculations, e.g. simulations in states at $\lambda=0$ and $\lambda=1$ may be initialized with the independently generated ligand coordinates. Another example would be using true apo and holo protein structures as the end states to initialize simulations for absolute binding free energy calculations. 
%Bhati \textit{et al.} recently published two studies directly comparing the equilibrium (EQ) and NEQ approaches at unprecedented scales for both RBFE and ABFE calculations using their ensemble simulation-based TIES method~\cite{wan2023eqvsneq, bhati2025}. Further, they systematically studied the effect of various parameters involved in setting up calculations for both approaches and recommended optimal protocols in each case for industrial applications. They highlighted several limitations of the NEQ approach that should be accounted for while performing such calculations to ensure reliability of predictions.

@@

    \item \href{http://www.ties-service.org}{TIES} allows the setup of perturbation maps, hybrid topologies and structures as well as simulation-ready models for NAMD and OpenMM \url{https://github.com/UCL-CCS/TIES} \cite{bieniek2023}.

@@

	\item Alchemlyb: Multipackage free energy analysis
	\url{https://github.com/alchemistry/alchemlyb}~\cite{daviddotson2020alchemistry}.
	\item pymbar: MBAR implementation, but have to roll your own analysis wrapper      
	\url{https://github.com/choderalab/pymbar} \cite{shirts2008statisticallya}.
	\item Arsenic: Standardising alchemical free energy analysis \url{https://github.com/openforcefield/Arsenic}
	\item Free Energy Workflows: Sire-specific free energy map analysis using weighted path averages \url{https://github.com/michellab/freenrgworkflows}.
    \item \href{http://www.ties-service.org}{TIES}: Perform statistical analysis of the output of alchemical ensemble simulations and report reliable predictions along with statistical uncertainties for a variety of free energy estimators \url{https://github.com/UCL-CCS/TIES} \cite{bieniek2023}.
	\end{itemize}
\item[] Generally, commercial software will offer more complete pipelines in which standalone analysis applications are not necessarily needed; free and open source packages often require manual analysis but allow more flexibility and modification.
\end{itemize}
